### 2020-01-15 analysis/OEDcheck.Rmd

---
title: "curiobaby_drop OED"
author: "George Kachergis"
date: "1/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
require(tidyverse)
```

## Load data

```{r, load-data}
dirs = list.files("agent-data")
d = data.frame()
for(dir in dirs) {
  filenames <- list.files(paste("agent-data",dir,sep='/'), full.names=TRUE)
  for(f in filenames) {
    #tmp <- read.csv(f, header=T)
    tmp <- read.table(f, sep=',', header=T, stringsAsFactors = F)
    d = rbind(d, tmp)
  }
}
```

`r nrow(d)` simulations with `r length(unique(d$dropped_objects))` unique dropped objects, of which we only used a subset of 9 for our behavioral experiments. We filter the simulations with undesired objects (sphere, cylinder, and cube).

```{r, count-unique-subsets}
names(d) # collision_count

hash_cols <- function(dat, cols) {
  shapes = c("Bowl", "Cylinder", "Cone", "Cube", "Dumbbell", "Octahedron", 
             "Pentagon", "Pipe", "Pyramid", "Ring", "Sphere", "TriangularPrism")
  for(i in 1:length(cols)) {
    dat[,cols[i]] = as.numeric(factor(dat[,cols[i]], levels=shapes)) #* 10^(i-1)
  }
  for(r in 1:nrow(dat)) {
    dat[r,cols] = sort(dat[r,cols])
  }
  for(i in 1:length(cols)) {
    dat[,cols[i]] = dat[,cols[i]] * 10^(i-1)
  }
  hash = rowSums(dat[,cols])
  return(hash)
}

unwanted = c("Sphere","Cylinder","Cube")

d$drop_set = hash_cols(d, 8:10)
d$ground_set = hash_cols(d, 5:7)

df <- d %>% 
  filter(
    !ground_object_1 %in% unwanted,
    !ground_object_2 %in% unwanted,
    !ground_object_3 %in% unwanted,
    !available_drop_object_1 %in% unwanted,
    !available_drop_object_2 %in% unwanted,
    !available_drop_object_3 %in% unwanted
  ) # 3200 left

length(unique(d$drop_set)) # 285 unique drop sets...
length(unique(d$ground_set)) # 285 unique ground sets..

A = c("Pyramid", "Torus", "TriangularPrism")
B = c("Cone", "Octahedron", "Pipe")
C = c("Bowl", "Dumbbell", "Pentagon")
d$DropSet = NA
for(r in 1:nrow(d)) {
  model_drop_objs = d[r,paste0("available_drop_object_", 1:3)]
  d[r,]$DropSet = ifelse(setequal(model_drop_objs, A), "A", 
                    ifelse(setequal(model_drop_objs, B), "B",
                      ifelse(setequal(model_drop_objs, C), "C", NA)))
}

xp = na.omit(d)

#d = subset(d, episode>299) # Michael says first 300 are warm-up off-policy
d = read_csv("agent/20191203-084115/agent_0.csv") # most recent = trained dynamics model

# selected drop object is in collision_object_1 or _2, so target is case where dropped_object!=collision_object
d <- d %>% mutate(
  collision_object_1 = ifelse(collision_object_1=="None", NA, collision_object_1),
  collision_object_2 = ifelse(collision_object_2=="None", NA, collision_object_2),
  target = ifelse(collision_object_1==dropped_object, collision_object_2, collision_object_1)
  )

```

## Drop Preferences

```{r}

drops = sort(table(d$dropped_object), decr=T) # 13 objects

round(drops / sum(drops), 2)

# should normalize by available_drop_objects_1 _2 and _3
```

Some changing preference in drop objects: cube, pentagon, torus, and cone > octahedron and bowl
Thirteen objects here, but we've only printed 9 (though could add a sphere easily, at least)

## Target Preferences

Are objects roughly equally available as targets?

```{r}
all_ground = c(d$ground_object_1, d$ground_object_2, d$ground_object_3)
sort(round(table(all_ground) / sum(table(all_ground)), 2), decr=T)
```

Roughly...

```{r}
# sometimes collision_object_1 == collision_object_2 == None ?
targ = sort(table(d$target), decr=T)
round(targ / sum(targ), 2)
```

Collision count:
```{r}
round(table(d$collision_count) / sum(table(d$collision_count)), 2)
```