### 2020-02-02 analysis/model_analysis.Rmd
model data cleanup and inspection

---
title: "curiobaby_drop model analysis"
author: "George Kachergis"
date: "1/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
require(tidyverse)
```

## Load data

```{r, load-data}
load_model_data <- function(data_dir, policy_name) {
  dirs = paste("../analysis/agent-data", data_dir, sep='/')
  d = data.frame()
  for(dir in dirs) {
    subdirs <- list.files(dir, full.names=TRUE)
    for(sub in subdirs) {
      filenames <- list.files(sub, full.names=TRUE)
      for(f in filenames) {
        #tmp <- read.csv(f, header=T)
        tmp <- read.table(f, sep=',', header=T, stringsAsFactors = F)
        d = rbind(d, tmp)
      }
    }
  }
  d$policy = policy_name
  return(d)
}

cur = load_model_data("curiodrop_01_curious", "curious")
noncur = load_model_data("curiodrop_01_noncurious", "noncurious")
rand = load_model_data("random_policy", "random")
all = rbind(cur, noncur, rand)
names(all) # collision_count drop_target

drop_loc_cols = c("ground_object_1_distance_on_floor_contact",
                  "ground_object_2_distance_on_floor_contact",
                  "ground_object_3_distance_on_floor_contact")
all$ClosestObjDistance = apply(all[,drop_loc_cols], 1, min) # distance of closest object

# TargetSpace is object that is closest
md <- all %>% mutate(TargetSpace = case_when(
        ClosestObjDistance==ground_object_1_distance_on_floor_contact ~ ground_object_1,
        ClosestObjDistance==ground_object_2_distance_on_floor_contact ~ ground_object_2,
        ClosestObjDistance==ground_object_3_distance_on_floor_contact ~ ground_object_3
      )) %>% mutate(TargetSpace = case_when(
        TargetSpace=="RDumbbell" ~ "Dumbbell",
        TRUE ~ TargetSpace
      )) %>% mutate(DropChoice = case_when(
        dropped_object=="RDumbbell" ~ "Dumbbell",
        TRUE ~ dropped_object
      )) %>% mutate(DropLocation = case_when(
        drop_target=="floor" ~ "space",
        drop_target=="RDumbbell" ~ "Dumbbell",
        TRUE ~ drop_target
      ))

# create StimSet column 
md <- md %>% mutate(DropSet = case_when(
        drop_object_partitioning=="Bowl_Pentagon_RDumbbell" ~ "c",
        drop_object_partitioning=="Pyramid_Torus_TriangularPrism" ~ "a",
        drop_object_partitioning=="Cone_Octahedron_Pipe" ~ "b"
      )) %>% mutate(TargetSet = case_when(
        ground_object_partitioning=="Bowl_Pentagon_RDumbbell" ~ "c",
        ground_object_partitioning=="Pyramid_Torus_TriangularPrism" ~ "a",
        ground_object_partitioning=="Cone_Octahedron_Pipe" ~ "b"
      )) %>% mutate(StimSet = paste0(TargetSet, DropSet))

table(md$policy)

md <- md %>% select("episode", "policy", "StimSet", "DropChoice", "DropLocation", "TargetSpace", "DropSet", "TargetSet")

# episodes with same drop and target sets (does model show pref to drop objs on same objs like pilot children?)
same_dt = subset(md, DropSet==TargetSet)

# don't want these due to our pilot results (kids show same pref)
md <- md %>% filter(DropSet!=TargetSet)
save(md, file="agent-data/model_data.RData")
```

`r nrow(all)` simulations with the 9 desired objects. We filter the `r nrow(same_dt)` episodes that had the same drop set and target set, as this trial type was not used in the behavioral experiment.

```{r, collisions-and-distance}

hist(all$collision_count)
hist(all$ClosestObjDistance)
weird = subset(all, ClosestObjDistance<0) # 1715 episodes where the distance to closest object is -1 
table(weird$hit_object_before_ground) # FALSE for all 
table(weird$drop_target) # mostly floor, but some not
```

## Drop Preferences

```{r}

drops = table(md$policy, md$DropChoice)

round(drops / rowSums(drops), 2)

```

Collision count:
```{r}
round(table(all$collision_count) / sum(table(all$collision_count)), 2)
```