### 2021-06-03 analysis/exp3_analysis.Rmd
exp3 analysis start

---
title: 'Experiment 3: ratings'
author: "George"
date: "6/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
library(png)
library(grid)
library(ggplot2)
library(xtable)
require(here)
require(tidyverse)
require(tidyboot)
library(apa) # for generating APA-style text to report statistical tests
library(ggpubr)
require(gridExtra)
require(papaja)
require(kableExtra)
```

## Load Data

```{r pressure, echo=FALSE}

load(file=here("paper/processed_data.RData"))

int_dat <- read_csv(here("analysis/data_exp3/curiodrop V2 data - Intestingness.csv")) 
lik_dat <- read_csv(here("analysis/data_exp3/curiodrop V2 data - Likelihood.csv"))

```

## Clean data

```{r}
install.packages("qualtRics")
```