### 2021-06-03 analysis/exp3_analysis.Rmd
making summary tables

---
title: 'Experiment 3: ratings'
author: "George"
date: "6/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
library(png)
library(grid)
library(ggplot2)
library(xtable)
require(here)
require(tidyverse)
require(tidyboot)
library(apa) # for generating APA-style text to report statistical tests
library(ggpubr)
require(gridExtra)
require(papaja)
require(kableExtra)
require(qualtRics)
```

## Load Data

```{r pressure, echo=FALSE}
# Preregistration: https://docs.google.com/document/d/10iQ80L4K9PSa8yINPcjURj7xl10hXxzFHqtsS3DnsII/edit
load(file=here("paper/processed_data.RData"))

int_dat <- read_csv(here("analysis/data_exp3/curiodrop V2 data - Intestingness.csv")) %>% 
  select(-`c1-14a`, -`c2-21a`, -`c3-22a`, -`c4-28a`, -`Q9`, -`Q8_1`)
lik_dat <- read_csv(here("analysis/data_exp3/curiodrop V2 data - Likelihood.csv")) %>%
  select(-`c1-14a`, -`c2-21a`, -`c3-22a`, -`c4-28a`, -`Q9`, -`Q8_1`)

new_col_names = unlist(int_dat[1,])
new_col_names[1:2] = c("Exclusion","SC0")
new_col_names[length(new_col_names)] = "subjID"
names(int_dat) = new_col_names
names(lik_dat) = new_col_names

int_dat = int_dat[2:nrow(int_dat),]
lik_dat = lik_dat[2:nrow(lik_dat),]
```

## Clean data

```{r}
# ToDo: count exclusions
int_dat <- int_dat %>% filter(is.na(Exclusion))

lik_dat <- lik_dat %>% filter(is.na(Exclusion))
```

# Pivot

```{r}
int_long <- int_dat %>% relocate(subjID) %>% # SC0 = score of catch trials
  pivot_longer(cols=4:ncol(int_dat), names_to="trial", values_to = "rating") %>%
  separate(col=trial, sep='-', into=c("drop","choice")) %>% select(-Exclusion) %>%
  rename(catch_trials_correct=`SC0`)

lik_long <- lik_dat %>% relocate(subjID) %>%
  pivot_longer(cols=4:ncol(int_dat), names_to="trial", values_to = "rating") %>%
  separate(col=trial, sep='-', into=c("drop","choice")) %>% select(-Exclusion) %>%
  filter(!is.na(rating))
```

## Data inspection

```{r}
table(int_long$subjID)
table(lik_long$subjID)

sum(is.na(int_long$rating))
sum(is.na(lik_long$rating))
```

## Summarize Choice Data

```{r}
adult_agg <- adult_long %>% group_by(drop, target, relation) %>%
  summarize(chose_target = mean(chose_target)) # do we want raw counts for chose target vs. foil?
# for trial 1, 70% chose target (30% chose distractor)

child_agg <- child_long %>% group_by(drop, target, relation) %>%
  summarize(chose_target = mean(chose_target))
```

## Summarize Ratings Data

```{r}
lik_agg <- lik_long %>% mutate(rating = as.numeric(rating)) %>%
  group_by(drop, choice) %>%
  summarise(likelihood = mean(rating))

int_agg <- lik_long %>% mutate(rating = as.numeric(rating)) %>%
  group_by(drop, choice) %>%
  summarise(interest = mean(rating)) 
# also want ratings scaled 0-1?
```